---
title: "Introduction to Rtpca"
author:
- name: Nils Kurzawa
  affiliation: 
  - European Molecular Biology Laboratory (EMBL), Genome Biology Unit
date: "`r format(Sys.time(), '%d %B, %Y')`"
package: Rtpca
output:
    BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{Introduction to Rtpca}
    %\VignetteEngine{knitr::rmarkdown}
    %VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

`Rtpca` 

This vignette only represents a minimal example. To have a look at a more realistic example feel free to check out:  

```{r, message=FALSE, warning=FALSE}
library(TPP)
library(Rtpca)
```

## Import Thermal proteome profiling data using the TPP package
```{r}
data("hdacTR_smallExample")
```

Filter hdacTR_data to speed up computations
```{r}
set.seed(123)
random_proteins <- sample(hdacTR_data[[1]]$gene_name, 300)
```

```{r}
hdacTR_data_fil <- lapply(hdacTR_data, function(temp_df){
    filter(temp_df, gene_name %in% random_proteins)
})
```

We can now import our small example dataset using the import function from the `TPP` package:
```{r}
trData <- tpptrImport(configTable = hdacTR_config, data = hdacTR_data_fil)
```

```{r}
data("ori_et_al_complex_ppis")
```

## Run TPCA on data from a single condition

We can run TPCA for protein-protein interactions like this:
```{r}
vehTPCA <- runTPCA(
    objList = trData,
    ppiAnno = ori_et_al_complex_ppis
)
```
We can also run TPCA to test for coaggregation of protein complexes like this:
```{r}
data("ori_et_al_complexes_df")

vehComplexTPCA <- runTPCA(
    objList = trData,
    complexAnno = ori_et_al_complexes_df,
    minCount = 2
)
```

We can plot a ROC curve for how well our data captures protein-protein interactions:
```{r}
plotPPiRoc(vehTPCA, computeAUC = TRUE)
```

And we can also plot a ROC curve for how well our data captures protein complexes:

```{r}
plotComplexRoc(vehComplexTPCA, computeAUC = TRUE)

```


## Run differential TPCA on two conditions 

In order to test for protein-protein interactions that change significantly between both conditions, we can run the `runDiffTPCA` as illustrated below:

```{r}
diffTPCA <- 
    runDiffTPCA(
        objList = trData[1:2], 
        contrastList = trData[3:4],
        ppiAnno = ori_et_al_complex_ppis)
```

We can then plot a volcano plot to visualize the results:
```{r}
plotDiffTpcaVolcano(diffTPCA)
```

As mentioned above, this vignette includes only a very minimal example, have a look at a more extensive example [here](https://github.com/nkurzaw/Rtpca_analysis/blob/master/Hashimoto_et_al_analysis.pdf).

```{r}
sessionInfo()
```

